\documentclass[hyperref, a4paper]{article}

\usepackage{geometry}
\usepackage{titling}
\usepackage{titlesec}
% No longer needed, since we will use enumitem package
% \usepackage{paralist}
\usepackage{enumitem}
\usepackage{footnote}
\usepackage{enumerate}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{physics}
\usepackage{tensor}
\usepackage{siunitx}
\usepackage[version=4]{mhchem}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{autobreak}
\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{nameref,zref-xr}
\zxrsetup{toltxlabel}
\zexternaldocument*[hw2-]{../2/2}[2.pdf]
\usepackage[colorlinks,unicode]{hyperref} % , linkcolor=black, anchorcolor=black, citecolor=black, urlcolor=black, filecolor=black
\usepackage[most]{tcolorbox}
\usepackage{prettyref}

% Page style
\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}
\titlespacing{\paragraph}{0pt}{1pt}{10pt}[20pt]
\setlength{\droptitle}{-5em}
\preauthor{\vspace{-10pt}\begin{center}}
\postauthor{\par\end{center}}

% More compact lists 
\setlist[itemize]{
    itemindent=17pt, 
    leftmargin=1pt,
    listparindent=\parindent,
    parsep=0pt,
}

% Math operators
\DeclareMathOperator{\timeorder}{\mathcal{T}}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\legpoly}{P}
\DeclareMathOperator{\primevalue}{P}
\DeclareMathOperator{\sgn}{sgn}
\newcommand*{\ii}{\mathrm{i}}
\newcommand*{\ee}{\mathrm{e}}
\newcommand*{\const}{\mathrm{const}}
\newcommand*{\suchthat}{\quad \text{s.t.} \quad}
\newcommand*{\argmin}{\arg\min}
\newcommand*{\argmax}{\arg\max}
\newcommand*{\normalorder}[1]{: #1 :}
\newcommand*{\pair}[1]{\langle #1 \rangle}
\newcommand*{\fd}[1]{\mathcal{D} #1}
\DeclareMathOperator{\bigO}{\mathcal{O}}

% TikZ setting
\usetikzlibrary{arrows,shapes,positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.markings}
\tikzstyle arrowstyle=[scale=1]
\tikzstyle directed=[postaction={decorate,decoration={markings,
    mark=at position .5 with {\arrow[arrowstyle]{stealth}}}}]
\tikzstyle ray=[directed, thick]
\tikzstyle dot=[anchor=base,fill,circle,inner sep=1pt]

% Algorithm setting
% Julia-style code
\SetKwIF{If}{ElseIf}{Else}{if}{}{elseif}{else}{end}
\SetKwFor{For}{for}{}{end}
\SetKwFor{While}{while}{}{end}
\SetKwProg{Function}{function}{}{end}
\SetArgSty{textnormal}

\newcommand*{\concept}[1]{{\textbf{#1}}}

% Embedded codes
\lstset{basicstyle=\ttfamily,
  showstringspaces=false,
  commentstyle=\color{gray},
  keywordstyle=\color{blue}
}

% Reference formatting
\newrefformat{fig}{Figure~\ref{#1} on page~\pageref{#1}}

% Color boxes
\tcbuselibrary{skins, breakable, theorems}
\newtcbtheorem[number within=section]{warning}{Warning}%
  {colback=orange!5,colframe=orange!65,fonttitle=\bfseries, breakable}{warn}
\newtcbtheorem[number within=section]{note}{Note}%
  {colback=green!5,colframe=green!65,fonttitle=\bfseries, breakable}{note}
\newtcbtheorem[number within=section]{info}{Info}%
  {colback=blue!5,colframe=blue!65,fonttitle=\bfseries, breakable}{info}

\newcommand{\hwtwo}{\href{../2/2.pdf}{Homework 2}}

\title{QFT I, Homework 4}
\author{Jinyuan Wu}

\begin{document}

\maketitle

\paragraph{Scalar QED} Consider the theory of a complex scalar field $\phi$ interacting with the electromagnetic field $A^{\mu}$. The Lagrangian is
\begin{equation}
    \mathcal{L}=-\frac{1}{4} F_{\mu \nu} F^{\mu \nu}+\left(D_{\mu} \phi\right)^{*} D^{\mu} \phi-m^{2} \phi^{*} \phi.
    \label{eq:scalar-qed}
\end{equation}
where $D_{\mu}=\partial_{\mu}+ \ii e A_{\mu}$ is the usual gauge covaraint derivative.
\begin{itemize}
    \item[(a)] Show the Lagrangian is invariant under the gauge transformations
    \begin{equation}
        \phi(x) \rightarrow \ee^{-\ii \alpha(x)} \phi(x), \quad A_{\mu}(x) \rightarrow A_{\mu}(x)+\frac{1}{e} \partial_{\mu} \alpha(x).
        \label{eq:gauge}
    \end{equation}
    \item[(b)] Derive the Feynman rules for the interaction between photons and scalar particles.
    \item[(c)] Draw all the leading-order Feynman diagrams and compute the amplitude for the process $\gamma \gamma \rightarrow \phi \phi^{*}$.
    \item[(d)] Compute the differential cross section $\dd \sigma / \dd \cos \theta$. You can take an average over all initial state polarizations. For simplicity, you can restrict your calculation in the limit $m=0$.
    \item[(e)] Draw all leading order Feynman diagrams, that contribute to the Compton scattering process $\gamma \phi \rightarrow \gamma \phi$ and compute the differential cross section $\dd \sigma / \dd \cos \theta$ with $m=0$.
\end{itemize}

\paragraph{Solution} \begin{itemize}
\item[(a)] Under the gauge transformation \eqref{eq:gauge}, we have 
\[
    F_{\mu \nu}  \to 
    F'_{\mu \nu} = \partial_\mu A'_\nu - \partial_\nu A'_\mu
    = \partial_\mu \left(A_\nu + \frac{1}{e} \partial_\nu \alpha\right) 
    - \partial_\nu \left(A_\mu + \frac{1}{e} \partial_\mu \alpha\right)
    = \partial_\mu A_\nu - \partial_\nu A_\mu = F_{\mu \nu},
\] 
so the first term in \eqref{eq:scalar-qed} remains the same. It is obvious that under \eqref{eq:gauge}
\[
    \phi^* \phi \to \phi'^* \phi' = \ee^{\ii \alpha} \phi^* \ee^{- \ii \alpha} \phi = \phi^* \phi,
\]
so the third term in \eqref{eq:scalar-qed} is also invariant. Also we have
\[
    \begin{aligned}
        D^\mu \phi \to (\partial^\mu + \ii e A'^\mu) \phi' 
        &= (\partial^\mu + \ii e A^\mu + \ii \partial^\mu \alpha) \ee^{- \ii \alpha} \phi  \\
        &= \ee^{- \ii \alpha} (\partial^\mu - \ii \partial^\mu \alpha 
        + \ii e A^\mu + \ii \partial^\mu \alpha) \phi \\
        &= \ee^{- \ii \alpha} D^\mu \phi,
    \end{aligned}
\]
and also 
\[
    (D^\mu \phi)^* = \ee^{\ii \alpha} D^\mu \phi, 
\]
so $D^\mu \phi (D^\mu \phi)^*$ is also invariant. 
Therefore \eqref{eq:scalar-qed} is invariant under \eqref{eq:gauge}.

\item[(b)] We make the following expansion of Fourier transformation. For the complex scalar field we have 
\begin{equation}
    \phi(x) = \int \frac{\dd[3]{\vb*{p}}}{(2\pi)^3} \frac{1}{\sqrt{2 \omega_{\vb*{p}}}} (a_{\vb*{p}} \ee^{- \ii p \cdot x} + b^\dagger_{\vb*{p}} \ee^{\ii p \cdot x}).
\end{equation}
which was proved in \eqref{hw2-eq:complex-scalar-expansion} in \hwtwo. The vector field is expanded as 
\begin{equation}
    A_\mu(x) = \int \frac{\dd[3]{\vb*{p}}}{(2\pi)^3} \frac{1}{\sqrt{2 \omega_{\vb*{p}}}} \sum_{r=1}^2 \epsilon_\mu^r(\vb*{p}) \left({a}_{\vb*{p}, r}^\dagger \ee^{ \ii p \cdot x} + {a}_{\vb*{p}, r} \ee^{ - \ii p \cdot x} \right).
\end{equation}

Expanding \eqref{eq:gauge} we have 
\begin{equation}
    \mathcal{L} = \mathcal{L}_\text{scalar} + \mathcal{L}_\text{vector} + \mathcal{L}_\text{scalarQED},
\end{equation} 
where $\mathcal{L}_\text{scalar}$ and $\mathcal{L}_\text{vector}$ are Lagrangians of free scalar field and 
free massless vector field, and 
\begin{equation}
    \begin{aligned}
        \mathcal{L}_\text{scalarQED} &= (D_{\mu} \phi)^{*} D^{\mu} \phi - (\partial_\mu \phi)^* \partial^\mu \phi \\
        &= e^2 \eta_{\mu \nu} A^\mu A^\nu \phi^* \phi - \ii e A_\mu \phi^* \partial^\mu \phi + \ii e \partial_\mu \phi^* A^\mu \phi .
    \end{aligned}
\end{equation}
The first term has no derivatives. Therefore it gives the following (momentum space) vertex:
\input{vertex-1.tex}
where the factor $\ii$ comes from the time evolution operator and the factor $2$ comes from the fact that there 
are two identical photon lines. The two $\phi$ lines can be any of the following four:
\input{lines.tex}
The second term gives
\[
    - \ii e A_\mu \phi^* \partial^\mu\phi \sim - \ii e A_{\mu} 
    (a^\dagger_{\vb*{p}} \ee^{\ii p \cdot x} + b_{\vb*{p}} \ee^{- \ii p \cdot x}) 
    (- \ii (p' \cdot x) a_{\vb*{p}'} \ee^{- \ii p' \cdot x} + \ii (p' \cdot x) b^\dagger_{\vb*{p}'} \ee^{\ii p' \cdot x}),
\]
and the third term is its complex conjugate. Therefore, the $a^\dagger a$ term in the Lagrangian is 
\[
    \sim - e(p_1 + p_2)_\mu A^\mu a^\dagger_{\vb*{p}_1} a_{\vb*{p}_2},
\]
so after adding the $\ii$ factor from the time evolution operator we have  
\input{vertex-2.tex}
and we can change the direction of a momentum line and a $\phi$-particle line arbitrarily; if a momentum line 
goes in contrast to the corresponding particle line, then we need to add a minus sign to the corresponding 
momentum. For example we have
\input{vertex-3.tex}
There are four vertices in this type in total.

\begin{note*}{}{}
    Here we follow the notation of Peskin, i.e. using the \emph{momentum} arrow to denote whether this line
    represents creation or annihilation and using the arrow \emph{on} a particle line to show whether this line 
    represents a particle (if the direction of the particle line is parallel to the direction of the momentum line)
    or a antiparticle (otherwise). The real direction of a 4-momentum is \emph{not} represented in any arrow.
\end{note*} 

\item[(c)] We enumerate over all possible diagrams. The vertex \eqref{eq:vertex-1} itself is a diagram:
\input{diagram-gg-1.tex} 
Combining two \eqref{eq:vertex-2}-type vertices we have a $t$-channel
\input{diagram-gg-3.tex}
and a $u$-channel
\input{diagram-gg-2.tex}

\begin{note*}{}{}
    We \emph{do not} need to distinguish the direction of the $q'$ momentum line. This line can be either 
    a particle line or an antiparticle line, but since the ordinary propagator $\ii / (p^2 - m^2 + \ii 0^+)$ 
    is obtained by summing up the two cases, when we write down this propagator, we have automatically 
    considered both processes. 
\end{note*}

Summing everything up, we have 
\begin{equation}
    \begin{aligned}
        \ii \mathcal{M}(\gamma \gamma \to \phi \phi^*) &= \ii (\mathcal{M}_4 + \mathcal{M}_t + \mathcal{M}_u) \\
        &= \ii e^2 (\epsilon^\sigma)^\mu (\epsilon^{\sigma'})^\nu \left( 2 \eta_{\mu \nu} + \frac{(k-2p)_\mu (k'-2q)_\nu}{t - m^2} + \frac{(k-2q)_\mu (k'-2p)_\nu}{u - m^2} \right) \\
        &\eqqcolon \ii (\epsilon^\sigma)^\mu (\epsilon^{\sigma'})^\nu \mathcal{M}_{\mu \nu} ,
    \end{aligned}
\end{equation}
where 
\begin{equation}
    t = (k-p)^2, \quad u = (q-k)^2.
\end{equation}

\item[(d)] We work in the center-of-mass frame, and therefore we have $k = (\abs*{\vb*{k}}, \vb*{k})$, 
and $k' = (\abs*{\vb*{k}}, - \vb*{k})$. 
The massless limit can be calculated with Eq.~(4.85) in Peskin, which is 
\begin{equation}
    \left(\dv{\sigma}{\Omega}\right)_\text{CM} = \frac{\abs*{\mathcal{M}}^2}{64 \pi^2 E_\text{CM}^2},
\end{equation} 
What we need is $\abs*{\mathcal{M}}^2$. We have 
\[
    \begin{aligned}
        \abs*{\mathcal{M}}^2 &= \sum_{\epsilon^\sigma, \epsilon^{\sigma'}} P(\epsilon^\sigma, \epsilon^{\sigma'})
        (\epsilon^\sigma)^\mu (\epsilon^{\sigma'})^\nu (\epsilon^\sigma)^{\rho *} (\epsilon^{\sigma'})^{\delta *} 
        \mathcal{M}_{\mu \nu} \mathcal{M}_{\rho \delta}^* \\
        &= \frac{1}{4} \sum_{\sigma= \pm 1}  (\epsilon^\sigma)^\mu (\epsilon^\sigma)^{\rho *}
        \sum_{\sigma' = \pm 1} (\epsilon^{\sigma'})^\nu (\epsilon^{\sigma'})^{\delta *}
        \mathcal{M}_{\mu \nu} \mathcal{M}_{\rho \delta}^* .
    \end{aligned}
\]
It can be verified that in a coordinate where $\vb*{p}_1$ is on the $\pm \vu*{z}$ direction, we have 
\[
    \sum_\sigma (\epsilon^\sigma)^\mu (\epsilon^\sigma)^{\rho *} = \diag(0, 1, 1, 0),
\] 
and in our case, it can be verified that 
\begin{equation}
    \sum_\sigma (\epsilon^\sigma)^\mu (\epsilon^\sigma)^{\rho *} = - \eta^{\mu \rho} + 
    \frac{k^\mu k'^\rho + k^\rho k'^\mu }{k \cdot k'},
\end{equation}
and a similar result holds for the sum on $\sigma'$, which can be obtained by replacing all the indices. 
We therefore have 
\[
    \begin{aligned}
        \sum_\sigma (\epsilon^\sigma)^\mu (\epsilon^\sigma)^{\rho *} \mathcal{M}_{\mu \rho} &= e^2
        \Big( - 2 \delta^\rho_\nu - \frac{(k-2p)^\rho (k'-2q)_\nu}{t} - \frac{(k-2q)^\rho (k'-2p)_\nu}{u} 
        + 2 \frac{k_\nu k'^\rho + k'_\nu k^\rho}{k \cdot k'} \\
        &\quad \quad + \frac{k \cdot (k-2p) k'^\rho (k'-2q)_\nu + k' \cdot (k - 2p) k^\rho (k'-2q)_\nu}{(k \cdot k') t} \\
        &\quad \quad + \frac{k \cdot (k-2q) k'^\rho (k'-2p)_\nu + k' \cdot (k - 2q) k^\rho (k'-2p)_\nu}{(k \cdot k') u} \Big),
    \end{aligned}
\]
and similarly 
\[
    \begin{aligned}
        \sum_{\sigma'} (\epsilon^{\sigma'})^\nu (\epsilon^{\sigma'})^{\delta *} \mathcal{M}_{\rho \delta}^* &=
        e^2 \Big( - 2 \delta_\rho^\nu - \frac{(k-2p)_\rho (k'-2q)^\nu}{t} - \frac{(k-2q)_\rho (k'-2p)^\nu}{u} 
        + 2 \frac{k^\nu k'_\rho + k'^\nu k_\rho}{k \cdot k'} \\
        &\quad \quad + \frac{k' \cdot (k'-2q) k^\nu (k-2p)_\rho + k \cdot (k' - 2q) k'^\nu (k-2p)_\rho}{(k \cdot k') t} \\
        &\quad \quad + \frac{k' \cdot (k'-2p) k^\nu (k-2q)_\rho + k \cdot (k' - 2p) k'^\nu (k-2q)_\rho}{(k \cdot k') u} \Big).
    \end{aligned}
\]
Multiplication of the two expressions, after expansion, is 

The process is too long to be displayed here. The equation above has been simplified using the fact 
that $k^2 = k'^2 = p^2 = q^2 = 0$.

\end{itemize}

\end{document}