\documentclass[hyperref, a4paper]{article}

\usepackage{geometry}
\usepackage{titling}
\usepackage{titlesec}
% No longer needed, since we will use enumitem package
% \usepackage{paralist}
\usepackage{enumitem}
\usepackage{footnote}
\usepackage{marginnote}
\usepackage{enumerate}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{physics}
\usepackage{tensor}
\usepackage{siunitx}
\usepackage[version=4]{mhchem}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{autobreak}
\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{nameref,zref-xr}
\zxrsetup{toltxlabel}
\zexternaldocument*[solid-]{../solid/solid}[solid.pdf]
\zexternaldocument*[optics-]{../optics/optics}[optics.pdf]
\usepackage[colorlinks,unicode]{hyperref} % , linkcolor=black, anchorcolor=black, citecolor=black, urlcolor=black, filecolor=black
\usepackage[most]{tcolorbox}
\usepackage{prettyref}

% Page style
\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}
\titlespacing{\paragraph}{0pt}{1pt}{10pt}[20pt]
\setlength{\droptitle}{-5em}
\preauthor{\vspace{-10pt}\begin{center}}
\postauthor{\par\end{center}}

% More compact lists 
\setlist[itemize]{
    itemindent=17pt, 
    leftmargin=1pt,
    listparindent=\parindent,
    parsep=0pt,
}

% Math operators
\DeclareMathOperator{\timeorder}{\mathcal{T}}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\legpoly}{P}
\DeclareMathOperator{\primevalue}{P}
\DeclareMathOperator{\sgn}{sgn}
\newcommand*{\ii}{\mathrm{i}}
\newcommand*{\ee}{\mathrm{e}}
\newcommand*{\const}{\mathrm{const}}
\newcommand*{\suchthat}{\quad \text{s.t.} \quad}
\newcommand*{\argmin}{\arg\min}
\newcommand*{\argmax}{\arg\max}
\newcommand*{\normalorder}[1]{: #1 :}
\newcommand*{\pair}[1]{\langle #1 \rangle}
\newcommand*{\fd}[1]{\mathcal{D} #1}
\DeclareMathOperator{\bigO}{\mathcal{O}}

% TikZ setting
\usetikzlibrary{arrows,shapes,positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.markings}
\tikzstyle arrowstyle=[scale=1]
\tikzstyle directed=[postaction={decorate,decoration={markings,
    mark=at position .5 with {\arrow[arrowstyle]{stealth}}}}]
\tikzstyle ray=[directed, thick]
\tikzstyle dot=[anchor=base,fill,circle,inner sep=1pt]

% Algorithm setting
% Julia-style code
\SetKwIF{If}{ElseIf}{Else}{if}{}{elseif}{else}{end}
\SetKwFor{For}{for}{}{end}
\SetKwFor{While}{while}{}{end}
\SetKwProg{Function}{function}{}{end}
\SetArgSty{textnormal}

\newcommand*{\concept}[1]{{\textbf{#1}}}

% Embedded codes
\lstset{basicstyle=\ttfamily,
  showstringspaces=false,
  commentstyle=\color{gray},
  keywordstyle=\color{blue}
}

% Reference formatting
\newrefformat{fig}{Figure~\ref{#1} on page~\pageref{#1}}

% Color boxes
\tcbuselibrary{skins, breakable, theorems}
\newtcbtheorem[number within=section]{warning}{Warning}%
  {colback=orange!5,colframe=orange!65,fonttitle=\bfseries, breakable}{warn}
\newtcbtheorem[number within=section]{note}{Note}%
  {colback=green!5,colframe=green!65,fonttitle=\bfseries, breakable}{note}
\newtcbtheorem[number within=section]{info}{Info}%
  {colback=blue!5,colframe=blue!65,fonttitle=\bfseries, breakable}{info}

\newcommand{\soliddoc}{\href{../solid/solid.pdf}{this solid state physics note}}
\newcommand{\opticsdoc}{\href{../optics/optics.pdf}{this optics note}}

\title{Electron Gas with Coulomb Interaction}
\author{Jinyuan Wu}

\begin{document}

\maketitle

It is said that for a simple demonstration of what happens in a metal, we can only work with the \concept{jellium model},
ignoring the details of the lattice, which means we can just work with a non-relativistic electron gas with Coulomb
interaction. \marginnote{Zhengzhong Li, Sec.~4.2} This article is an extension of Section~\ref{solid-sec:high-density} in \soliddoc. We will discuss some early development of RPA.

\section{Notations and basic facts about the jellium model} 

We define 
\begin{equation}
    \rho(\vb*{r}) = \sum_\sigma \psi^\dagger_\sigma(\vb*{r}) \psi_\sigma(\vb*{r}) 
    = \frac{1}{V} \sum_{\vb*{k}, \vb*{k}', \sigma} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k}' \sigma} \ee^{- \ii (\vb*{k} - \vb*{k}') \cdot \vb*{r}} 
    = \frac{1}{V} \sum_{\vb*{q}} \sum_{\vb*{k}, \sigma} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} + \vb*{q}, \sigma} \ee^{\ii \vb*{q} \cdot \vb*{r}},
\end{equation}
and the commutation relations are 
\begin{equation}
    \comm*{c_{\vb*{k}}}{c^\dagger_{\vb*{k}'}} = \delta_{\vb*{k} \vb*{k}'}.
\end{equation}
We can also write down $\rho(\vb*{r})$ in the first quantization formulation, i.e. 
\begin{equation}
    \rho(\vb*{r}) = \sum_i \delta(\vb*{r} - \vb*{r}_i) = \sum_i \frac{1}{V} \sum_{\vb*{q}} \ee^{\ii \vb*{q} \cdot (\vb*{r} - \vb*{r}_i)} = \frac{1}{V} \sum_{\vb*{q}} \sum_i \ee^{\ii \vb*{q} \cdot (\vb*{r} - \vb*{r}_i)}.
\end{equation}
We have 
\begin{equation}
    \rho_{\vb*{q}} \coloneqq \sum_i \ee^{- \ii \vb*{q} \cdot \vb*{r}_i} = \sum_{\vb*{k}, \sigma} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} + \vb*{q}, \sigma} , \quad 
    \rho(\vb*{r}) = \frac{1}{V} \sum_{\vb*{q}} \rho_{\vb*{q}} \ee^{\ii \vb*{q} \cdot \vb*{r}}.
    \label{eq:density}
\end{equation}
Easily, we find 
\begin{equation}
    \rho_{\vb*{q} = 0} = \int \dd[3]{\vb*{r}} \rho(\vb*{r}) = N.
\end{equation}
The Hamiltonian of electrons is 
\[
    H_\text{electron} = \sum_{\vb*{k}, \sigma} \epsilon_{\vb*{k}} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} \sigma} 
    + \frac{1}{2V} \sum_{\vb*{k}, \vb*{k}', \vb*{q} , \sigma, \sigma'} 
    c^\dagger_{\vb*{k} + \vb*{q}, \sigma}  c^\dagger_{\vb*{k}' - \vb*{q}, \sigma'} V(q)
    c_{\vb*{k}' \sigma'} c_{\vb*{k} \sigma},
\]
where 
\begin{equation}
    V(q)= \int \ee^{- \ii \vb*{q} \cdot \vb*{r}} \frac{e^2}{r} = \frac{4 \pi e^2}{\vb*{q}^2}.
\end{equation}
The Hamiltonian of the lattice and the electron-lattice coupling is 
\[
    H_\text{lattice} = \frac{1}{2V} \sum_{\vb*{q}} \rho_\text{ion,$-\vb*{q}$} V(q)\rho_\text{ion,$\vb*{q}$} -
    \frac{1}{V} \sum_{\vb*{q}} \rho_\text{ion,$\vb*{q}$} V(q) \rho_{\vb*{q}}.
\]
When $\abs*{\vb*{q}} = 0$, $V(q)$ is divergent, but this does not matter. 
In the jellium model, the only non-zero Fourier component of $\rho_\text{ion}(\vb*{r})$ is 
$\rho_\text{ion,$\vb*{q}=0$} = N$ (the solid is neutral so the number of positive charges must be equal to 
the number of negative charges), and we soon find 
\[
    H_\text{lattice} + \frac{1}{2V} \sum_{\vb*{q} = 0, \vb*{k}, \vb*{k}', \sigma, \sigma'} c^\dagger_{\vb*{k} + \vb*{q}, \sigma}  c^\dagger_{\vb*{k}' - \vb*{q}, \sigma'} V(q)
    c_{\vb*{k}' \sigma'} c_{\vb*{k} \sigma} = 0.
\]
So all divergences cancel with each other, and in the end, the total Hamiltonian is 
\begin{equation}
    \begin{aligned}
        H &= \sum_{\vb*{k}, \sigma} \epsilon_{\vb*{k}} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} \sigma} 
        + \frac{1}{2V} \sum_{\vb*{q} \neq 0} \sum_{\vb*{k}, \vb*{k}', \sigma, \sigma'} c^\dagger_{\vb*{k} + \vb*{q}, \sigma}  c^\dagger_{\vb*{k}' - \vb*{q}, \sigma'} V(q)
        c_{\vb*{k}' \sigma'} c_{\vb*{k} \sigma} \\
        &= \sum_{\vb*{k}, \sigma} \epsilon_{\vb*{k}} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} \sigma} 
        + \frac{1}{2V} \sum_{\vb*{q} \neq 0} \rho^\dagger_{\vb*{q}} V(q) \rho_{\vb*{q}} 
        = \sum_{\vb*{k}, \sigma} \epsilon_{\vb*{k}} c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} \sigma} 
        + \frac{1}{2V} \sum_{\vb*{q} \neq 0} \rho_{-\vb*{q}} V(q) \rho_{\vb*{q}}.
    \end{aligned}
    \label{eq:whole-ham}
\end{equation}
The contribution of the positive ion ``jell'' both constrains the electrons in the solid (or in other words,
give a chemical potential) and regularize the singularity of $V(\vb*{q} = 0)$. 

Note that \eqref{eq:whole-ham} differs with 
\begin{equation}
    H = \sum_{i} \frac{\vb*{p}_i^2}{2m} + \frac{1}{2} \sum_{i \neq j} \frac{e^2}{\abs*{\vb*{r}_i - \vb*{r}_j}}
\end{equation} 
with the $i=j$ term. This is an infinite constant and does not matter. 

\section{Classical (or first quantized) theory of the collective oscillation of electrons} 

From \eqref{eq:density}, we have \marginnote{Zhengzhong Li, Sec.~4.3}
\[
    \dot{\rho}_{\vb*{q}} = \sum_j (- \ii \vb*{q}) \cdot \vb*{v}_j \ee^{- \ii \vb*{q} \cdot \vb*{r}_j},
\]
\begin{equation}
    \ddot{\rho}_{\vb*{q}} = \sum_j (- \ii \vb*{q} \cdot \dot{\vb*{v}}_j - (\vb*{q} \cdot \vb*{v}_j)^2) \ee^{- \ii \vb*{q} \cdot \vb*{r}_j}.
    \label{eq:rho-q-eom}
\end{equation}
Now $\dot{\vb*{v}}_j$ can be derived from \eqref{eq:whole-ham}:
\[
    \begin{aligned}
        m \dot{\vb*{v}}_j &= - \grad_j \frac{1}{2V} \sum_{\vb*{q} \neq 0} V(q) \rho^\dagger_{\vb*{q}} \rho_{\vb*{q}} \\
        &= - \frac{1}{2V} \grad_j \sum_{\vb*{q} \neq 0} V(q) \sum_{i, k} \ee^{\ii \vb*{q} \cdot (\vb*{r}_i - \vb*{r}_k)} \\
        &= - \frac{1}{2V} \sum_{\vb*{q} \neq 0} V(q) \sum_k (\ii \vb*{q}) \ee^{\ii \vb*{q} \cdot (\vb*{r}_j - \vb*{r}_k)} + \sum_i (- \ii \vb*{q}) \ee^{\ii \vb*{q} \cdot (\vb*{r}_j - \vb*{r}_i)} \\
        &= - \frac{1}{V} \sum_{\vb*{q} \neq 0} \ii \vb*{q} V(q) \sum_i \ee^{\ii \vb*{q} \cdot (\vb*{r}_j - \vb*{r}_i)} \\
        &= - \frac{4 \pi e^2}{V} \sum_{\vb*{q} \neq 0} \frac{\ii \vb*{q}}{q^2} \ee^{\ii \vb*{q} \cdot \vb*{r}_j} \rho_{\vb*{q}},
    \end{aligned}
\]
and from \eqref{eq:rho-q-eom}, we have 
\[
    \begin{aligned}
        \ddot{\rho}_{\vb*{q}} &= - \sum_j \frac{4 \pi e^2}{m V} \sum_{\vb*{q}' \neq 0} \frac{\vb*{q} \cdot \vb*{q}'}{q'^2} \ee^{\ii \vb*{q}' \cdot \vb*{r}_j} \rho_{\vb*{q}'} \ee^{- \ii \vb*{q} \cdot \vb*{r}_j} - \sum_j (\vb*{q} \cdot \vb*{v}_j)^2 \ee^{- \ii \vb*{q} \cdot \vb*{r}_j} \\
        &= - \frac{4 \pi e^2}{m V} \sum_{\vb*{q}' \neq 0} \rho_{\vb*{q}'} \rho_{\vb*{q} - \vb*{q}'} - \sum_j (\vb*{q} \cdot \vb*{v}_j)^2 \ee^{- \ii \vb*{q} \cdot \vb*{r}_j}.
    \end{aligned}
\]
Now we make the \concept{random phase approximation (RPA)} in its original form: We assume that only the 
$\vb*{q} = \vb*{q}'$ term in the first term is important, because in the high density limit, there are no 
position preference of electrons (when the density is low, there might be a Wigner crystal, and RPA fails), 
and when $\vb*{q} \neq 0$, both $\rho_{\vb*{q}'}$ and $\rho_{\vb*{q} - \vb*{q}'}$ are sums of almost random 
phase factors $\ee^{- \ii \vb*{q} \cdot \vb*{r}_j}$, and therefore are both small enough. So we get the EOM 
after RPA:
\begin{equation}
    \begin{aligned}
        \ddot{\rho}_{\vb*{q}} &= - \frac{4 \pi e^2}{mV} \rho_{\vb*{q}} \rho_{0} - \sum_j (\vb*{q} \cdot \vb*{v}_j)^2 \ee^{- \ii \vb*{q} \cdot \vb*{r}_j} \\
        &= - \frac{4 \pi e^2 n}{m} \rho_{\vb*{q}} - \sum_j (\vb*{q} \cdot \vb*{v}_j)^2 \ee^{- \ii \vb*{q} \cdot \vb*{r}_j} ,
    \end{aligned}
\end{equation}  
where $n = N / V$ is the jellium density. Section~\eqref{solid-sec:linear-response-energy-band} in \soliddoc{} 
tells us that the electron-hole pair excitations are gapless, but from the EOM above we soon find that in the 
$\vb*{q} \to 0$ case there is a finite $\omega$ solution, which is given by 
\begin{equation}
    \ddot{\rho}_{\vb*{q}} + \omega_\text{p}^2 \rho_{\vb*{q}} = 0, \quad \omega_\text{p}^2 = \frac{4 \pi e^2 n}{m}.
\end{equation}
We see that this term arises from the Coulomb interaction. In other words, long range interaction induces a 
collective modes in the metal, which is now known as \concept{plasmon}. We know it is plasmon, or quantized 
plasma oscillation, because our derivation above also works for the oscillation of negative charges around 
positive charges in a plasma.

\section{Second quantized EOM of density modes}

Now we try to derive plasmon in the second quantization EOM framework. Generally speaking, a generalized \marginnote{Zhengzhong Li, Sec.~4.4}
hydrodynamic mode in an electron gas is made of a linear combination of 
\begin{equation}
    \rho_{\vb*{k} \vb*{q}}^\dagger \eqqcolon \sum_\sigma c^\dagger_{\vb*{k} + \vb*{q}, \sigma} c_{\vb*{k} \sigma}.
\end{equation}
By evaluating the EOM, we may be able to identify stable density modes.

We start from the free theory. It can be verified that 
\begin{equation}
    \comm*{c_1^\dagger c_2}{c_3^\dagger c_4} = \delta_{23} c_1^\dagger c_4 - \delta_{14} c_3^\dagger c_2,
\end{equation} 
and therefore we have 
\[
    \begin{aligned}
        \ii \dot{\rho}_{\vb*{k} \vb*{q}}^\dagger &= \comm*{{\rho}_{\vb*{k} \vb*{q}}^\dagger}{H_0} 
        = \sum_{\vb*{p}, \alpha} \frac{\vb*{p}^2}{2m} \sum_\sigma \comm*{c^\dagger_{\vb*{k} + \vb*{q}, \sigma} c_{\vb*{k} \sigma}}{c^\dagger_{\vb*{p} \alpha} c_{\vb*{p} \alpha}} \\
        &= \sum_{\vb*{p}, \alpha, \sigma} \frac{\vb*{p}^2}{2m} (c^\dagger_{\vb*{k} + \vb*{q}, \sigma} c_{\vb*{p} \alpha} \delta_{\vb*{k} \vb*{p}} \delta_{\sigma \alpha} - c^\dagger_{\vb*{p} \alpha} c_{\vb*{k} \sigma} \delta_{\vb*{k}+\vb*{q}, \vb*{p}} \delta_{\sigma \alpha}) \\
        &= \sum_{\sigma} \left( \frac{\vb*{k}^2}{2m} c^\dagger_{\vb*{k}+\vb*{q}, \sigma} c_{\vb*{k} \sigma} - \frac{(\vb*{q} + \vb*{k})^2}{2m} c^\dagger_{\vb*{k}+\vb*{q}, \sigma} c_{\vb*{k} \sigma} \right),
    \end{aligned}
\]
so 
\begin{equation}
    \ii \dot{\rho}^\dagger_{\vb*{k} \vb*{q}} = \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{H_0} = - \omega^\text{pair}_{\vb*{k} \vb*{q}} {\rho}^\dagger_{\vb*{k} \vb*{q}}, 
    \quad \omega^\text{pair}_{\vb*{k} \vb*{q}} = \frac{(\vb*{q} + \vb*{k})^2}{2m} - \frac{\vb*{k}^2}{2m}.
\end{equation}
Therefore, we find there are electron-hole pairs in the free theory. In the last section we find that in a plasmon
mode, we have $\rho_{\vb*{k}} = \sum_{\vb*{q}} \rho_{\vb*{k} \vb*{q}} \sim \ee^{- \ii \omega_{\vb*{p}} t}$, 
and $\omega_{\vb*{q} = 0} \neq 0$, but here when $\vb*{q} = 0$, $\omega^\text{pair}
_{\vb*{k} \vb*{q}}$ vanishes, so there is no plasmon. This is expected, since in the last section we find 
the plasmon comes from the Coulomb interaction.

Now we move on to discuss the jellium model. We have 
\[
    \begin{aligned}
        \comm*{\rho^\dagger_{\vb*{k} \vb*{q}}}{H} &= \comm*{\rho^\dagger_{\vb*{k} \vb*{q}}}{H_0 + H_\text{Coulomb}} = 
        - \omega^\text{pair}_{\vb*{k} \vb*{q}} {\rho}^\dagger_{\vb*{k} \vb*{q}} + \frac{1}{2V} \sum_{\vb*{q}' \neq 0} V(q') \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{\rho_{-\vb*{q}'} \rho_{\vb*{q}'}} \\
        &= - \omega^\text{pair}_{\vb*{k} \vb*{q}} {\rho}^\dagger_{\vb*{k} \vb*{q}} + \frac{1}{2V} \sum_{\vb*{q}' \neq 0} V(q') (\rho_{-\vb*{q}'} \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{ \rho_{\vb*{q}'}} + \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{ \rho_{-\vb*{q}'}} \rho_{\vb*{q}'} ) \\
        &= - \omega^\text{pair}_{\vb*{k} \vb*{q}} {\rho}^\dagger_{\vb*{k} \vb*{q}} + \frac{1}{2V} \sum_{\vb*{q}' \neq 0} V(q') (\rho_{-\vb*{q}'} \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{ \rho_{\vb*{q}'}} + \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{ \rho_{\vb*{q}'}} \rho_{-\vb*{q}'} ) ,
    \end{aligned}
\] 
where 
\[
    \begin{aligned}
        \comm*{\rho_{\vb*{k} \vb*{q}}^\dagger}{ \rho_{\vb*{q}'}} &= 
        \sum_{\alpha} \sum_{\vb*{k}', \beta} \comm*{c_{\vb*{k}+\vb*{q}, \alpha}^\dagger c_{\vb*{k} \alpha}}{c^\dagger_{\vb*{k}' \beta} c_{\vb*{k}' + \vb*{q}', \beta}} \\
        &= \sum_{\alpha} ( c^\dagger_{\vb*{k} + \vb*{q}, \alpha} c_{\vb*{k} + \vb*{q}', \alpha} - c^\dagger_{\vb*{k} + \vb*{q} - \vb*{q}' , \alpha} c_{\vb*{k} \alpha} ).
    \end{aligned}
\]
So the EOM of $\rho_{\vb*{k} \vb*{q}}^\dagger$ is finally  \marginnote{Zhengzhong Li (4.4.9)}
\begin{equation}
    \ii \dot{\rho}_{\vb*{k} \vb*{q}}^\dagger = - \omega^\text{pair}_{\vb*{k} \vb*{q}} {\rho}^\dagger_{\vb*{k} \vb*{q}} - \frac{1}{2V} \sum_{\vb*{q}' \neq 0, \alpha} V(q') \acomm*{\rho_{-\vb*{q}'}}{c^\dagger_{\vb*{k} + \vb*{q} - \vb*{q}' , \alpha} c_{\vb*{k} \alpha} - c^\dagger_{\vb*{k} + \vb*{q}, \alpha} c_{\vb*{k} + \vb*{q}', \alpha}}.
    \label{eq:bosonic-eq}
\end{equation}

By taking all possible momenta in \eqref{eq:bosonic-eq}, we get a closed group of equations about density 
modes. In principle, this gives the (generalized) hydrodynamic modes in the jellium model completely, 
though we all know it is highly difficult to solve EOM of operators. This can also be seen as a successful 
\emph{bosonization} the theory, though unlike the case in a Luttinger liquid, this does not help us understand
the behavior of the jellium model. A possible way to simplify 
\eqref{eq:bosonic-eq} is to assume that $c^\dagger_{\vb*{k} + \vb*{q} - \vb*{q}' , \alpha} c_{\vb*{k} \alpha} 
- c^\dagger_{\vb*{k} + \vb*{q}, \alpha} c_{\vb*{k} + \vb*{q}', \alpha}$ does not fluctuate much, and we 
can replace it by the expectation value. if under this approximation we find a stable mode, 
then we have 
\[
    \begin{aligned}
        (\omega - \omega_{\vb*{k} \vb*{q}}^\text{pair}) \rho^\dagger_{\vb*{k} \vb*{q}} &= \frac{1}{2V} \sum_{\vb*{q}' \neq 0, \alpha} V(q') \acomm*{\rho_{-\vb*{q}'}}{\expval*{c^\dagger_{\vb*{k} + \vb*{q} - \vb*{q}' , \alpha} c_{\vb*{k} \alpha} - c^\dagger_{\vb*{k} + \vb*{q}, \alpha} c_{\vb*{k} + \vb*{q}', \alpha}}} \\
        &= \frac{1}{2V} \sum_{\vb*{q}' \neq 0, \alpha} 2 V(q') \rho_{\vb*{q}'}^\dagger (n_{\vb*{k}} \delta_{\vb*{q} \vb*{q}'} - n_{\vb*{k} + \vb*{q}} \delta_{\vb*{q} \vb*{q}'}) \\
        &= \frac{2}{V} \rho^\dagger_{\vb*{q}} (n_{\vb*{k}} - n_{\vb*{k} + \vb*{q}}) V(q),
    \end{aligned}
\] 
where 
\begin{equation}
    n_{\vb*{k}} \coloneqq \expval*{c^\dagger_{\vb*{k} \sigma} c_{\vb*{k} \sigma}}. \quad \quad (\text{no summation over $\sigma$})
\end{equation}
From the equation above we find 
\begin{equation}
    \rho_{\vb*{k} \vb*{q}}^\dagger = \frac{2 V(q)}{V} \rho^\dagger_{\vb*{q}} \frac{n_{\vb*{k}} - n_{\vb*{k} + \vb*{q}}}{\omega - \omega_{\vb*{k} \vb*{q}}^\text{pair}},
    \label{eq:rpa-operator-result}
\end{equation}
and summing over $\vb*{k}$, we have 
\begin{equation}
    1 = \frac{2 V(q)}{V} \sum_{\vb*{k}} \frac{n_{\vb*{k}} - n_{\vb*{k} + \vb*{q}}}{\omega - \omega_{\vb*{k} \vb*{q}}^\text{pair}}.
    \label{eq:freq-eq}
\end{equation}

What we are doing here is actually a mean field approximation. 
It is natural to guess that the approximation corresponds to one Green function resummation strategy, 
which we will discuss in detail in \prettyref{sec:green}.
An intuitive way to see what is going on is to note that \eqref{eq:rpa-operator-result} is actually
\input{rpa-diagram.tex}
where the $(n_{\vb*{k}} - n_{\vb*{k} + \vb*{q}}) / (\omega - \omega_{\vb*{k} \vb*{q}}^\text{pair})$
factor is ``the propagator of an electron-hole pair''. The approximation we have made is also 
a \emph{random phase approximation}, because after the approximation, all $\vb*{q}' \neq \vb*{q}$
terms in \eqref{eq:bosonic-eq} disappear. 

Now we can easily find the frequencies and the states of all density modes. This also shows a benefit 
of our operator EOM based approach: in a Green function based approach, we can only easily find the 
spectrum of the density modes by checking the singularities of $\expval*{nn}$, but it is not that 
easy to identify their actual ``shape''. We, however, will postpone the solution of \eqref{eq:freq-eq} \marginnote{}
to \prettyref{sec:boson-modes}, since TODO: can we find the lifetime of the excitons purely by this section's approach?

\section{The Green function theory}\label{sec:green}

We have already discussed the density-density Green function in the free theory in \eqref{solid-eq:ext-electron-retarded-green-function} in \soliddoc. 


\section{The electric susceptibility and bosonic modes}\label{sec:boson-modes}

Now it is time to evaluate the analytic properties of the susceptibility. \marginnote{Zhengzhong Li, Sec.~4.6}
The dispersion relations are given by $\epsilon = 0$ (not $\epsilon \to \infty$ - see 
Section~\ref{optics-sec:green-and-linear-response} in \opticsdoc). 
Note that the imaginary part of $\epsilon$ indicates damping (see Section~\ref{optics-sec:in-interior-uniform}
in \opticsdoc), and therefore if $\Im \epsilon \neq 0$, even if on an $(\omega, \vb*{k})$ point we have 
$\Re \epsilon = 0$, this is not an eigenmode. However, if $\Im \epsilon$ is not large, we can still say 
that ``there is a mode at $(\omega, \vb*{k})$ with a finite lifetime''. So what we are going to do is 
to solve the equation $\Re \epsilon(\omega, \vb*{k}) = 0$ and check the lifetime of each mode.

\end{document}