\documentclass[a4paper]{article}

\usepackage{geometry}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{abstract}
\usepackage{amsmath, amssymb}
\usepackage{qtree}
\usepackage{gb4e}
\usepackage[colorlinks, urlcolor=cyan]{hyperref}
\usepackage{prettyref}

\geometry{left=3.18cm,right=3.18cm,top=2.54cm,bottom=2.54cm}

\title{Laudate, Laudate Dominum}
\author{Jinyuan Wu}

\begin{document}

\maketitle

\section{The Text}

\emph{Laudate, Laudate Dominum} is a hymn by Christopher Walker, a famous Catholic composer, and is written for the new Cathedral of Our Lady of the Angels, Los Angeles, California \cite{ocp-preview}.
The music is played in several videos available on Youtube \cite{youtube-1,youtube-2}, in one of which it was used as the entrance song in a mass held by Pope Francis. 

The lyrics can be found in OCP's website \cite{ocp-preview}. Since the material is copyrighted, we do not put a full copy here. You can download a preview from the link above.
The lyrics is written in Latin, English and Spanish.

Here is the Latin part of the lyrics:
\begin{quotation}
    Laudate, laudate Dominum, omnes gentes, laudate Dominum.

    Exsultate, jubilate per annos Domini, omnes gentes.
\end{quotation}

\section{Translation}

The Latin part can be glossed as follows:
\begin{exe} 
    \sn 
    \gll Laudate,               laudate                Dominum,     omnes       gentes,       laudate                Dominum. \\
         praise.IMP.PRS.ACT.2PL praise.IMP.PRS.ACT.2PL Lord.SG.ACC every.PL.VOC people.PL.VOC praise.IMP.PRS.ACT.2PL Lord.SG.ACC \\
    \glt Let us praise, Let us praise the Lord, O all people, let us praise the Lord.
    \sn 
    \gll Exsultate,              jubilate                per      annos Domini,     omnes        gentes. \\
         rejoice.IMP.PRS.ACT.2PL cheer.IMP.PRS.ACT.2PL   through  time.PL.ACC       Lord.SG.ACC every.PL.VOC people.PL.VOC \\
    \glt Let us rejoice, let us in the Lord's time, O all people.
\end{exe}   

\bibliographystyle{plain}
\bibliography{laudate} 

\end{document}